---
output: github_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	echo = FALSE,
	fig.align = "center",
	fig.width = 12,
	fig.height = 8,
	out.width = "100%"
)
```

# Benchmarking compboost vs. mboost

```{r, results="asis"}
cat(readd(my.system))
```

This document was automatically created using `drake`. To recreate this document just source `drake_benchmark.R`. 

## Runtime Benchmark

As already mentioned, the benchmark was conducted using `batchtools`. In case of benchmarking the runtime, we executed each algorithm for each configuration five times. The height of the bars of the following plots corresponds to the median of these five evaluations. The drawn error-bar illustrates the maximal and minimal measured runtime. 

Since we are interested in tracking the performance while varying three different parameter we vary one and fix the others at a specific value. The three interesting parameters are the number of iterations, number of observations, and number of features. To run the benchmark on your own machine it is sufficient to execute the `execute_runtime_benchmark.R` script and submit the jobs by calling `submitJobs()`. For the benchmark we have just used one due to a very high memory allocation of some jobs crashes the system on multiple processes.

To access the raw results you need to load the registry:

```{r, echo=TRUE, eval=FALSE}
loadRegistry("benchmark/runtime/benchmark_files")
```

After preprocessing the raw data are stored into a `data.frame` where each row represents a job with instances like the elapsed time and the dimension of the simulated data:

```{r}
loadd(raw.runtime.benchmark.data)
raw.runtime.benchmark.data[sample(seq_len(nrow(raw.runtime.benchmark.data)), 10), ] %>%
  knitr::kable(row.names = FALSE)
```

The preprocessing is defined in the `drake_runtime_benchmark.R` script where `raw.runtime.benchmark.data` is created. This also applies to the following graphics.

```{r, eval=any(raw.runtime.benchmark.data$time == 0), echo=FALSE, results="asis"}
cat("For any of the following bars with a height of zero it was not possible to execute the algorithm with the corresponding specification.")
```

### Increasing Number of Iterations

While increasing the number of iterations we fixed the number of observations at 2000, and the number of feature at 1000. Under this configuration we achieve a 15 times faster fitting process with `compboost` compared to `mboost` in boosting linear base-learner. Nevertheless, `glmboost` is faster due to the internal structure of `glmboost`, which is due to that all base-learners can be fitted in one matrix multiplication. But this approach is not suitable for `compboost` since it does not fit into the object-oriented system we provide. This is due to the flexibility in specifying ordinary base-learner combination and not making the whole fitting process conditionally on the used base-learners. Nevertheless, using spline base-learner, `compboost` is about five times faster than `mboost` and `glmboost` (which is just a wrapper of the original `mboost` algorithm).

```{r}
grid.draw(readd(runtime.plot.iterations))
```

Note that the relative factor highly depends on the number of observations. This behavior is described above.

### Increasing Number of Base-Learner

For increasing the number of base-learners we get a equal behavior as for increasing the number of iterations.
```{r, eval=any(raw.runtime.benchmark.data$time == 0), echo=FALSE, results="asis"}
cat("Nevertheless, with `mboost` it was not able to conduct the boosting on 4000 features while `compboost` it was.")
```
For this experiments we fix the number of observations at 2000 and the number of iterations at 1500.

```{r}
grid.draw(readd(runtime.plot.ncols))
```

### Increasing Number of Observations

This may have the biggest effect on computation time since increasing the number of observations affects the allocated memory as well as the size of the internal matrix multiplications. 

For a smaller number of observations, `compboost` definitely outperforms `mboost`. The relative runtime behavior decreases with increasing the number of observations. This is due to the size of the matrix multiplications which gets more weight then the whole boiler code, like initializing base-learner, which is very fast with `compboost`. For that reason and the reason that matrix multiplication is also not that slow in `R`, `mboost` comes closer to the performance of `compboost`.

```{r}
grid.draw(readd(runtime.plot.nrows))
```


## Memory Benchmark

For the memory benchmark every second was measured how much RAM was used. This curve is then plotted for each algorithm. 

In the case of the spline base-learner, `compboost` and `mboost` uses sparse matrices, which significantly reduces computing time and memory requirements. In general, `compboost` is efficient here too. A small exception is `glmboost`, which can also save memory due to its special structure.

```{r}
grid.draw(readd(memory.plot))
```