---
title: "What is component-wise boosting?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{About component-wise boosting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  # fig.path = "Readme_files/"
)

library(compboost)
```


In order to provide high flexibility we decided to write the base implementation of `compboost` in `C++` to make use of object oriented programming. These `C++` classes can be used in `R` after exposing them as `S4` classes. To take away another layer of abstraction we decided to break the `S4` classes down to a few `R6` classes wrapping the original ones.

All in all, the class system of `compboost` is a mixture of raw exposed `S4` classes and convenience classes written in `R6`. As usual for object oriented programming, the classes works on a reference base. This introduction aims to show how these references work and can be accessed.

The two classes that are most affected from the reference system are the **Response** and **Data** classes since they hold and transform data.

## Response Class

The target variable is represented by an object that inherits from the `Response` class. Depending on the target type we like to have different transformations of the internally predicted scores. For instance, having a binary classification task the score $\hat{f}(x) \in \mathbb{R}$ is transformed to a $[0,1]$ scale by using the logistic function:

$$
\hat{\pi}(x) = \frac{1}{1 + \exp(-\hat{f}(x))}
$$

To show the how references work here we first define a `ResponseBinaryClassif` object. Therefore, we use the `mtcars` dataset and create a new binary target variable for *fast* $\text{qsec} < 17$ or *slow* $\text{qsec} \geq 17$ cars and create the response object:
```{r}
df = mtcars[, c("mpg", "disp", "hp", "drat", "wt")]
df$qsec_cat = ifelse(mtcars$qsec < 17, "fast", "slow")

obj_response = ResponseBinaryClassif$new("qsec_cat", "fast", df$qsec_cat)
obj_response
```

To access the underlying representation of the response class (here a binary variable) one can use `$getResponse()`. In the initialization of a new response object the prediction $\hat{f} \in \mathbb{R}$ is initialized with zeros. We can also use the response object to calculate the transformed predictions $\hat{\pi} \in [0,1]$:
```{r}
knitr::kable(head(data.frame(
  target = df$qsec_cat,
  target_representation = obj_response$getResponse(),
  prediction_initialization = obj_response$getPrediction(),
  prediction_transformed = obj_response$getPredictionTransform()
)))
```

In the case of binary classification we can also use the response object to calculate the predictions on a label basis by using a specified threshold $a$:
$$
\hat{y} = 1 \ \ \text{if} \ \ \hat{\pi}(x) \geq a
$$

The default threshold here is 0.5:
```{r}
obj_resopnse$getThreshold()
head(obj_response$getPredictionResponse())
```

By setting the threshold to 0.6 we observe now that each class is predicted as negative:
```{r}
obj_response$setThreshold(0.6)
head(obj_response$getPredictionResponse())
```

This behavior has nothing to do with references at the moment. But, just prediction a score of 0 for all observations is no good predictor. During the fitting of a component-wise boosting model these predictions are adjusted over and over again by the `Compboost` object. This is where the reference comes in:
```{r}
cboost = boostSplines(data = df, target = obj_response, loss = LossBinomial$new(),
  iterations = 2000L, trace = 500L)
```

Having again a look at the predictions shows that they are different than before without touching the response class. This is because of the response object that is passed to the `Compboost` object. During the fitting process the predictions of the response object are set by the model:

```{r}
knitr::kable(head(data.frame(
  target = df$qsec_cat,
  prediction = obj_response$getPrediction(),
  prediction_transformed = obj_response$getPredictionTransform(),
  prediction_response = obj_response$getPredictionResponse()
)))
```

